---
title: "Chapter 1"
author: "Jens Christian Wahl"
date: "3.7.2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Data Visualization with ggplot2

We will in this chapter focus on the dataset __mpg__, which contains data 38 models of cars. 
We start out by plotting the relationship between a car's engine size (_displ_) and a car's fuel efficiency (_hwy_).
```{r,include = T}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

## Exercises

### 3 
The _drv_ variable describe if the car had four wheel drive, front wheel drive or rear wheel drive

### 4 - Make a scatterplot of _hwy_ vs. _cyl_

```{r, eval = FALSE}
ggplot(data = mpg) + geom_point(x = cyl, y = hwy) 
```
This will not be useful due to the fact that _cyl_ is a categorical variable. 

## Aesthetic Mappings

We want to figure out what type of cars are in the lower right corner. 

```{r}
#ggplot(data = mpg) + geom_point(mapping = aes(x=displ,y=hwy, color = class)) 
ggplot(data = mpg, aes(color = class)) + geom_point(mapping = aes(x=displ,y=hwy))
```
As we can see, they are all two seaters (sport cars). We use the `color = class` argument to map a unique color to each type of car. Notice the difference if we do it outside the `aes` block: 

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ,y = hwy), color = "blue")
```
Thus, when putting the argument outside the `aes`, it does not give any information about the variables, but only the appearance of the `geom_point`.

## Exercises 

### 2 - Which variables of `mpg` are categorical? 

All variables that are of type `chr` are factors.

```{r}
glimpse(mpg)
```

### 3 - Map a continuous variable to `color, size` and `shape`. 

```{r, warning=FALSE}
#color
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = year))
#size
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = year))
#shape - Does not work for continuous variables
#ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = year))

#Why so many black dots - Because we only have two years, 1999 and 2008
mpg %>% arrange(year)
```

### 4 - What does he stroke asthetic d0? 

The `stroke` aesthetic controls the width of the shape. The shape must have a border to work. 
```{r}
ggplot(mpg) + geom_point(mapping = aes(displ,hwy), stroke = 3)
```


## Facets








