---
title: "Chapter_3"
author: "Jens Christian Wahl"
date: "30 august 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```


```{r}
flights <- flights
head(flights)
glimpse(flights)
```
## Filter rows with `filter()`

```{r}
flights %>% filter(month == 1, day %in% c(1,2))
```

## Exercises with `filter()`
### Find all flights that: 


```{r, results='hide'}
#Arrival delay if two or more hours
flights %>% filter(arr_delay > 120)

#Flew to Houston (IAH or HOU)
glimpse(flights %>% filter(dest %in% c("IAH","HOU")))

#Were operated by United, American or Delta (UA, AA, DL)
glimpse(flights %>% filter(carrier %in% c("UA", "AA", "DL")))

#Departed in summer
glimpse(flights %>% filter(month %in% 7:9))

#Arrived more than two hours late but did not leave late
glimpse(flights %>% filter(arr_delay >= 60 & dep_delay <= 0))

#Delayed by at least one hour, but made up over 30 minutes in flight
glimpse(flights %>% filter(dep_delay >= 60 & arr_delay <( dep_delay - 30)))

#Departed between midnight and 5 a.m

glimpse(flights %>% filter(dep_time == 2400 | dep_time <= 600))
```

### What does `between()` do? 

The same as `x >= left & x <= right`. We could use if for 
```{r, results = 'hide'}

flights %>% filter(between(month,7,9))

```
### How many flights have missing `dep_time`? 

```{r}
flights %>% filter(is.na(dep_time))
```


These are canceled flights, since `arr_time`is missing. 

##Exercises with `arrange()`

### Sort all missing values to the top

```{r}
flights %>% arrange(desc(is.na(dep_time)), dep_time)
```
Note that `is.na()` returns a logical vector and `desc(TRUE) = -2, desc(FALSE) = -1`. That's why this works.

### Sort `flights` to find the most delayed flights. Find the flights that left earliest. 

```{r, results = 'hide'}
#Most delayed
flights %>% arrange(desc(dep_delay))

#Left earliest
flights %>% arrange(dep_delay)


```
### Sort flights to find the fastest flight 

```{r}
flights %>% arrange(air_time) %>% 
  select(origin, dest, air_time) %>% 
  head()
```


### Which flights traveled the longest? Which traveled the shortest? 

```{r}
flights %>% arrange(desc(distance)) %>%
  select(origin, dest, distance) %>%
  head()

flights %>% arrange(distance) %>%
  select(origin, dest, distance) %>%
  head()
```

## Exercises for `select()`

### Ways to select `dep_time, dep_delay, arr_time, arr_delay`

```{r, eval = FALSE}
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)

flights %>% select(starts_with("dep_"), starts_with("arr_"))

flights %>% select(ends_with("time"), ends_with("delay"))
```
`ends_with` does not work well since it returns several variables we are not interested in.

### What does `one_of()`do? 

Select variables based on a character vector instead of unquoted variable names.
```{r}
flights %>% select(one_of(c("year", "month")))
```

### Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
flights %>% select(contains("TIME"))
```

`contains()` does not care about upper or lower case letters. This can be changed by adding the argument `contains("TIME", ingnore.case = F)`.

## Exercises for `mutate()`

```{r}
#test
flights_sml <- flights %>%
  select(year:day,
         ends_with("delay"),
         distance, 
         air_time)
```

