---
title: "Chapter_3"
author: "Jens Christian Wahl"
date: "30 august 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```


```{r}
flights <- flights
head(flights)
glimpse(flights)
```
## Filter rows with `filter()`

```{r}
flights %>% filter(month == 1, day %in% c(1,2))
```

## Exercises with `filter()`
#### Find all flights that: 


```{r, results='hide'}
#Arrival delay if two or more hours
flights %>% filter(arr_delay > 120)

#Flew to Houston (IAH or HOU)
glimpse(flights %>% filter(dest %in% c("IAH","HOU")))

#Were operated by United, American or Delta (UA, AA, DL)
glimpse(flights %>% filter(carrier %in% c("UA", "AA", "DL")))

#Departed in summer
glimpse(flights %>% filter(month %in% 7:9))

#Arrived more than two hours late but did not leave late
glimpse(flights %>% filter(arr_delay >= 60 & dep_delay <= 0))

#Delayed by at least one hour, but made up over 30 minutes in flight
glimpse(flights %>% filter(dep_delay >= 60 & arr_delay <( dep_delay - 30)))

#Departed between midnight and 5 a.m

glimpse(flights %>% filter(dep_time == 2400 | dep_time <= 600))
```

#### What does `between()` do? 

The same as `x >= left & x <= right`. We could use if for 
```{r, results = 'hide'}

flights %>% filter(between(month,7,9))

```
#### How many flights have missing `dep_time`? 

```{r}
flights %>% filter(is.na(dep_time))
```


These are canceled flights, since `arr_time`is missing. 

##Exercises with `arrange()`

#### Sort all missing values to the top

```{r}
flights %>% arrange(desc(is.na(dep_time)), dep_time)
```
Note that `is.na()` returns a logical vector and `desc(TRUE) = -2, desc(FALSE) = -1`. That's why this works.

#### Sort `flights` to find the most delayed flights. Find the flights that left earliest. 

```{r, results = 'hide'}
#Most delayed
flights %>% arrange(desc(dep_delay))

#Left earliest
flights %>% arrange(dep_delay)


```
#### Sort flights to find the fastest flight 

```{r}
flights %>% arrange(air_time) %>% 
  select(origin, dest, air_time) %>% 
  head()
```


#### Which flights traveled the longest? Which traveled the shortest? 

```{r}
flights %>% arrange(desc(distance)) %>%
  select(origin, dest, distance) %>%
  head()

flights %>% arrange(distance) %>%
  select(origin, dest, distance) %>%
  head()
```

## Exercises for `select()`

#### Ways to select `dep_time, dep_delay, arr_time, arr_delay`

```{r, eval = FALSE}
flights %>% select(dep_time, dep_delay, arr_time, arr_delay)

flights %>% select(starts_with("dep_"), starts_with("arr_"))

flights %>% select(ends_with("time"), ends_with("delay"))
```
`ends_with` does not work well since it returns several variables we are not interested in.

#### What does `one_of()`do? 

Select variables based on a character vector instead of unquoted variable names.
```{r}
flights %>% select(one_of(c("year", "month")))
```

#### Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

```{r}
flights %>% select(contains("TIME"))
```

`contains()` does not care about upper or lower case letters. This can be changed by adding the argument `contains("TIME", ingnore.case = F)`.

## Exercises for `mutate()`

```{r}
#test
flights_sml <- flights %>%
  select(year:day,
         ends_with("delay"),
         distance, 
         air_time)
```

#### Convert `dep_time` and `sched_dep_time` to a more convenient representation of numbers of minutes since midnight. 

```{r}
flights %>% transmute(dep_time = (dep_time %/% 100)*60 + dep_time %% 100,
                      sched_dep_time = (sched_dep_time %/% 100)*60 + dep_time %% 100)
```

#### Compare `air_time` with `arr_time - dep_time` 

```{r}
flights %>% transmute(air_time,
                      diff_time = arr_time - dep_time)
```

We observed a difference because `arr_time` and `dep_time` are not on the format of minutes since midnight. We can fix it in the following way: 

```{r}
time2mins <- function(t){
  #Trasforms timepoint to minuttes since midnight
  (t %/% 100 * 60 + t %% 100) %% 1440
}

flights %>% transmute(air_time,
                      arr_time = time2mins(arr_time),
                      dep_time = time2mins(dep_time),
                      air_time2 = arr_time - dep_time, 
                      diff = air_time - air_time2)
```
#### Compare `dep_time, sched_dep_time, dep_delay`

```{r}
flights_sml <- flights %>%
  select(dep_time, sched_dep_time, dep_delay)

flights_sml %>% transmute(dep_time = time2mins(dep_time),
                       sched_dep_time = time2mins(sched_dep_time),
                       dep_delay,
                       diff_time = dep_time - (sched_dep_time + dep_delay))


```
#### Find the 10 most delayed flights using a ranking function.

`min_rank` will give rank in an increasing order, i.e. the lowest value will have rank 1. `min_rank(desc())` changes the sign of the variable and thus gives rank in a decreasing order.

```{r}
flights_delayed <- flights %>% mutate(dep_delayed_rank = min_rank(desc(dep_delay)))
flights_delayed %>% filter(dep_delayed_rank <= 10) %>% arrange(dep_delayed_rank)

```



#### 1:3 + 1:10

R recycles the shortest vector.

## Excercises for `summarize()` 

#### Come up with another approach that will give you the same output as `not_cancelled %>% count(dest)` and `not_cancelled %>% count(tailnum, wt = distance)` (without using `count()`)

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% group_by(dest) %>% summarise(n = n()) %>% arrange(desc(n))

not_cancelled %>% group_by(tailnum) %>% summarize(n = sum(distance))


```

#### Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay?

```{r}
cancelled <- flights %>% 
  mutate(cancelled = is.na(dep_delay) | is.na(arr_delay)) %>% 
  group_by(year, month, day) %>% 
  summarise(prop_cancelled = mean(cancelled),
            avg_dep_delay = mean(dep_delay, na.rm = T))


cancelled %>% ggplot(aes(x = avg_dep_delay, y = prop_cancelled)) + 
  geom_point() + geom_smooth()
```
Days that have big delays also have a higher proportion on cancelled flights. 

#### Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about `flights %>% group_by(carrier, dest) %>% summarise(n())`)

```{r}
#Worst carriers

```

